---
title: "BRAVE Kids 16S Analysis"
author: "Alex McCumber"
date: "11/3/2020"
output: html_document
---

```{r, include=FALSE}
#load some libraries
library(dada2)
library(dplyr)
library(tibble)
library(purrr)
library(DECIPHER)
library(phyloseq)
library(ggplot2)
library(Biostrings)
library(ggtree)
library(ape)
library(stringr)
library(vegan)
library(magrittr)
library(ape)
library(philr)
library(glmnet)
library(DESeq2)
library(tidyverse)
library(cluster)
library(devtools)
library(decontam)
library(picante)
library(vegan)
library(phyloseq)
library(tidyr)
library(nlme)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(compositions)
#Need to download the ancom function before running
source("~/BRAVE_Kids/ancom_v2.1.R")
```

```{r}
dir.create("~/BRAVE_Kids/")
```


#Set filepaths and get sample names
```{r, Set filepaths}
path = "/sharedspace/BraveKids/"

list.files(path)

fnFs = sort(list.files(path, pattern="_R1_001.fastq", full.names = TRUE))
fnRs = sort(list.files(path, pattern="_R2_001.fastq", full.names = TRUE))

sample.names = sapply(strsplit(basename(fnFs),"_L001"), `[`,1)
```


```{r}
plotQualityProfile(fnFs[1:20],aggregate = TRUE)
```


```{r}
plotQualityProfile(fnRs[1:20], aggregate = TRUE)
```


```{r, set file path for where filtered reads will go}
filtFs <- file.path(path, "filtered", paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(path, "filtered", paste0(sample.names, "_R_filt.fastq.gz"))
```


```{bash}
gunzip /sharedspace/BraveKids/PCOV1001_S1_L001_R1_001.fastq.gz

head -n 20 /sharedspace/BraveKids/PCOV1001_S1_L001_R1_001.fastq 

gzip /sharedspace/BraveKids/PCOV1001_S1_L001_R1_001.fastq

```


#Filter and trim reads, using standard calls and trimming 20 off each end to remove primers
```{r}
#set multithread to False if you're running Windows

out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(240,220),
              maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
              compress=TRUE, multithread=T)
head(out)

as.data.frame(out) %>%
  mutate(., sum(reads.out)/sum(reads.in))

#Remove samples with 0 reads
filter.out = out %>%
  as.data.frame(.) %>%
  rownames_to_column('samples') %>%
  filter(., reads.out > 1) %>%
  column_to_rownames('samples')

filter.samples = out %>%
  as.data.frame(.) %>%
  rownames_to_column('samples') %>%
  filter(., reads.out > 1) %>%
  column_to_rownames('samples') %>%
  rownames(.) 

#Re-establish sample names now that some are gone
sample.names = sapply(strsplit(basename(filter.samples),"_L001"), `[`,1)

filtFs <- file.path(path, "filtered", paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(path, "filtered", paste0(sample.names, "_R_filt.fastq.gz"))
```


#Generate error function, dereplicating, and merging reads
```{r, include=false}
#set multithread to False if you're running Windows
errF = learnErrors(filtFs, multithread = T)
errR = learnErrors(filtRs, multithread = T)

dadaFs <- dada(filtFs, err=errF, multithread=TRUE, pool = T)
dadaRs <- dada(filtRs, err=errR, multithread=TRUE, pool = T)

mergers <- mergePairs(dadaFs, filtFs, dadaRs, filtRs, verbose=TRUE)
```

```{r}
plotErrors(errF, nominalQ=TRUE)
```


```{r, making a sequence table}
seqtab = makeSequenceTable(mergers)

table(nchar(getSequences(seqtab)))

#Too many reads of lengths outside of anticipated amplicon size so will be removing all reads that don't fall within 250 to 260 bp
ncol(seqtab)

seqtab2=seqtab[,nchar(colnames(seqtab)) %in% seq(240,260)]

saveRDS(seqtab, "~/BRAVE_Kids/seqtab.RDS")

seqtab.nochim = removeBimeraDenovo(seqtab2, method="consensus", multithread = T, verbose = TRUE)

#3546 ASVs left
saveRDS(seqtab.nochim, "~/BRAVE_Kids/seqtab_nochim.RDS")

getN <- function(x) sum(getUniques(x))
track <- cbind(filter.out, sapply(mergers, getN), rowSums(seqtab.nochim))
colnames(track) <- c("input", "filtered", "merged", "nonchim")
rownames(track) <- sample.names
head(track)
colSums(track)
```

#Run this if you need to download the silva dataset v138 for IDTaxa function
```{bash, include = false}
wget http://www2.decipher.codes/Classification/TrainingSets/SILVA_SSU_r138_2019.RData -P ./silvadb
```

```{r}
dna <- DNAStringSet(getSequences(seqtab.nochim)) # Create a DNAStringSet from the ASVs

load("~/BRAVE_Kids/silvadb/SILVA_SSU_r138_2019.RData")

ids <- IdTaxa(dna, trainingSet, strand="top", processors=12, verbose=FALSE)

ranks <- c("domain", "phylum", "class", "order", "family", "genus") # ranks of interest

# Convert the output object of class "Taxa" to a matrix analogous to the output from assignTaxonomy
taxid <- t(sapply(ids, function(x) {
        m <- match(ranks, x$rank)
        taxa <- x$taxon[m]
        taxa[startsWith(taxa, "unclassified_")] <- NA
        taxa
}))
colnames(taxid) <- ranks; rownames(taxid) <- getSequences(seqtab.nochim)
```

##Creation of tree, need to run through python 2 (I think, I haven't been able to get this to work when I have python3 installed)
```{bash, make a tree with SEPP using gg backbone tree}
wget  "https://raw.github.com/smirarab/sepp-refs/master/gg/sepp-package.tar.bz"

tar xvfj sepp-package.tar.bz

cd sepp-package/sepp

python setup.py config -c

python setup.py install --user
```

```{r, make list of sequences in fasta file for tree making}
uniques=getUniques(seqtab.nochim)

sequences=names(uniques)

uniquesToFasta(uniques,fout = "~/BRAVE_Kids/seqs.fasta", ids=sequences)
```

```{bash}
./sepp-package/run-sepp.sh seqs.fasta BRAVE
```

```{r, view tree file with ggtree}
seq.tree = read.tree(file = "~/BRAVE_Kids/BRAVE_placement.tog.relabelled.tre") %>%
  keep.tip(., sequences)
```


#Import metadata and make dataframe
```{r}
metad = read_csv("~/BRAVE_Kids/metadata_16sv2.csv")

rownames(metad) = str_remove_all(metad$brave_id, "-")

metaDF = subset(metad, rownames(metad) %in% gsub("_.*","",sample.names))

rownames(metaDF) = str_remove_all(metaDF$brave_id, "-")

metaDF$X1=NULL

rm(metad)
```

#Create phyloseq object
```{r}
#need to change seqtab.nochim to have same sample names as metadata
rownames(seqtab.nochim) = gsub("_.*","",rownames(seqtab.nochim))

ps = phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE), 
               sample_data(metaDF), 
               tax_table(taxid),
              phy_tree(seq.tree))

ps.Bact=subset_taxa(ps, domain=="Bacteria") 

saveRDS(ps.Bact, "~/BRAVE_Kids/ps.RDS")

rm(ps)
```

#Decontam 
```{r}
library("decontam")
contamdf.freq = isContaminant(ps.Bact, method="frequency", conc = "dna", threshold = 0.1)

hist(contamdf.freq$p)

table(contamdf.freq$contaminant)

#Identified 40 contaminant taxa in at 0.1
#remove contaminant taxa

ps.noncontam = prune_taxa(!contamdf.freq$contaminant, ps.Bact)

set.seed(42)
plot_frequency(ps.Bact, taxa_names(ps.Bact)[sample(which(contamdf.freq$contaminant),4)], conc="dna") +
    xlab("DNA Concentration")
```

```{r}
subset_samples(ps.noncontam, sample_data(ps.noncontam)$brave_id == "extractNEGctrl") %>%
  prune_species(speciesSums(.) > 0, .)

#54 ASVs in extractNEGctrl

subset_samples(ps.noncontam, sample_data(ps.noncontam)$brave_id == "PCRneg") %>%
  prune_species(speciesSums(.) > 0, .)

#2 PCR Neg Ctrl ASVs 
```

#Remove contaminants ID'ed by decontam and remove ctrl and non-study samples
```{r, mesesage = FALSE}
ps.noncontam = prune_taxa(!contamdf.freq$contaminant, ps.Bact) %>%
  subset_samples(., sample_data(ps.noncontam)$sex != "NA") %>%
  prune_samples(sample_sums(.)>=1000, .) %>%
  #transform_sample_counts(.,function(x) x/sum(x)) %>%
  prune_species(speciesSums(.) > 0, .) 

ps.noncontam

saveRDS(ps.noncontam, "~/BRAVE_Kids/psBRAVE.RDS")

# Number ASVs and store the exact sequences in the reference sequence slot - can access with refseq(ps)
dna <- Biostrings::DNAStringSet(taxa_names(ps.noncontam))
names(dna) <- taxa_names(ps.noncontam)
phy.brave <- merge_phyloseq(ps.noncontam, dna)
taxa_names(ps.noncontam) <- paste0(seq(ntaxa(ps.noncontam)))
ps.noncontam
remove(dna)
```

```{r}
ps.noncontam = readRDS("~/BRAVE_Kids/psBRAVE.RDS")

#Subset to NP only samples and remove Indeterminate COVID postiive/negative results
ps.NP = subset_samples(ps.noncontam, sample_data(ps.noncontam)$method_np == "NP" & sample_data(ps.noncontam)$corona != "Indeterminate") %>%
  prune_species(speciesSums(.) > 0, .) 

ps.NP.r = transform_sample_counts(ps.NP,function(x) x/sum(x)) 

saveRDS(ps.NP, "~/BRAVE_Kids/psNP.RDS")
```

```{r, include = FALSE}
# Function to make matrix usable for vegan from phyloseq object
vegan_otu <- function(physeq) {
    OTU <- otu_table(physeq)
    if (taxa_are_rows(OTU)) {
        OTU <- t(OTU)
    }
    return(as(OTU, "matrix"))
}
```

#Code for Figures 1A and 1B and evaluate differences in shannon and chao1 diversity/richness
```{r}
sampleDF = sampleDF %>%
    mutate(agebin = case_when(age < 2 ~ '0-2',
                              age < 5 ~ '3-5',
                              age < 8 ~ '6-8',
                              age < 11 ~ '9-11',
                              age < 14 ~ '12-14',
                              age < 17 ~ '15-17',
                              TRUE ~ '18-20'))

Shannon = estimate_richness(ps.NP, measures = "Shannon") %>%
  cbind(., sampleDF, by = "names") 

Shannon$agebin = factor(Shannon$agebin, levels = c('0-2','3-5','6-8','9-11','12-14','15-17','18-20'))

sampleDF = sampleDF %>%
    mutate(Group1 = case_when(Group == 'Negative' ~ 'Negative',
                              Group == 'PosResp' ~ 'Positve (respiratory SX)',
                              Group == 'Pos_NoResp' ~ 'Positive (no respiratory SX)'))

ggplot(Shannon, aes(x=agebin, y=Shannon)) +
  geom_point(aes(color=Group1), alpha = 0.3) +
  scale_color_manual(values=c("darkslategray4", "mediumorchid4","#BF812D")) +
  stat_summary(aes(y=Shannon, group = Group1, color = Group1), fun.y = mean, geom = "line")+
  theme_bw()+
  xlab("Age (years)")+
  ylab("Shannon diversity") + 
  labs(color = "SARS-CoV-2 status") +
  scale_y_continuous(limits = c(0,4)) + 
  theme(legend.position = "none")

Chao = estimate_richness(ps.NP, measures = "Chao1") %>%
  cbind(., sampleDF, by = "names") 

Chao$agebin = factor(Chao$agebin, levels = c('0-2','3-5','6-8','9-11','12-14','15-17','18-20'))

ggplot(Chao, aes(x=agebin, y=Chao1)) +
  geom_point(aes(color=Group1), alpha = 0.3) +
  scale_color_manual(values=c("darkslategray4", "mediumorchid4","#BF812D")) +
  stat_summary(aes(y=Chao1, group = Group1, color = Group1), fun.y = mean, geom = "line")+
  theme_bw()+
  xlab("Age (years)")+
  ylab("Chao1 richness") + 
  labs(color = "SARS-CoV-2 status") +
  scale_y_continuous(limits = c(0,250)) + 
  theme(legend.position = "none")

summary(Chao$Chao1)

summary(Shannon$Shannon)

shapiro.test(Shannon$Shannon)

model = lm(Shannon ~ age, data = Shannon)
summary(model)

model = aov(Shannon ~ corona + age, Shannon)
summary(model)

#chao values are not normally distributed, so need to perform transformation
shapiro.test(Chao$Chao1)

Chao$logChao = log(Chao$Chao1)

#log chao1 values are normally distributed 
shapiro.test(Chao$logChao) #p = 0.33

model = lm(logChao ~ age+Group, data = Chao)
summary(model)

fit = aov(logChao ~ corona+age, data = Chao)
summary(fit)

model = aov(logChao ~ Group*age, data = Chao)
summary(model)
```

#Code for Figure 1C
```{r}
set.seed(6)

ord = ordinate(ps.NP, method = "NMDS", distance = "bray", trymax = 1000)

#converged at run 467

ord$stress # 0.238

sampleDF = sample_data(ps.NP.r) %>%
  data.frame(.)

sampleDF$agebin = factor(sampleDF$agebin, levels = c('0-2','3-5','6-8','9-11','12-14','15-17','18-20'))

sampleDF = sampleDF %>%
    mutate(facet = case_when(agebin == '0-2' ~ '0-2 years',
                              agebin == '3-5' ~ '3-5 years',
                              agebin == '6-8' ~ '6-8 years',
                              agebin == '9-11' ~ '9-11 years',
                              agebin == '12-14' ~ '12-14 years',
                              agebin == '15-17' ~ '15-17 years',
                              TRUE ~ '18-20 years'))

sampleDF$facet = factor(sampleDF$facet, levels = c('0-2 years','3-5 years','6-8 years','9-11 years','12-14 years','15-17 years','18-20 years'))

sampleDF = sampleDF %>%
    mutate(Group1 = case_when(Group == 'Negative' ~ 'Negative',
                              Group == 'PosResp' ~ 'Positve (respiratory symptoms)',
                              Group == 'Pos_NoResp' ~ 'Positive (no respiratory symptoms)'))

sample_data(ps.NP)$Group1 = sampleDF$Group1

ordDF = cbind(sampleDF, ord$points)
ordDF2=ordDF
ordDF$agebin=NULL
ordDF$facet=NULL

p=ggplot(data = ordDF, aes(x = MDS1, y = MDS2)) +
  geom_point(data=ordDF, alpha=0.05) + 
  geom_point(data = ordDF2, aes(color = Group1), alpha=0.7) + 
  facet_wrap(~ facet, ncol=4) +
  scale_color_manual(values = c("darkslategray4", "mediumorchid4","#BF812D")) + 
  theme_bw() +
  xlab("NMDS1") + 
  ylab("NMDS2") +
  labs(color = "SARS-CoV-2 status")+ 
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1))

PermDF = data.frame(sample_data(ps.NP))

BetaD=vegdist(vegan_otu(ps.NP), "bray")

#permanova by age
adonis(BetaD~age, permutations = 999, data=PermDF)

anosim(BetaD, PermDF$agebin, permutations = 999, distance = "bray",
    parallel = getOption("mc.cores"))

#look at age + covid (positive/negative) status
PermDF = data.frame(sample_data(ps.NP))

BetaD=vegdist(vegan_otu(ps.NP), "bray")

adonis(BetaD~age+corona, permutations = 999, data=PermDF)

```

#Code to generate Figure 2
```{r}
#create age bins and lots of long code to make mean abundance profiles

sampleDF = sample_data(ps.NP.r) %>%
  data.frame(.)

sampleDF = sampleDF %>%
    mutate(agebin = case_when(age < 3 ~ '0-2',
                              age < 6 ~ '3-5',
                              age < 9 ~ '6-8',
                              age < 12 ~ '9-11',
                              age < 15 ~ '12-14',
                              age < 18 ~ '15-17',
                              TRUE ~ '18-20'))

sample_data(ps.NP.r)$agebin = sampleDF$agebin

ps.Genus = tax_glom(ps.NP.r, taxrank = "genus") 

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$agebin == '0-2')

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$domain = NULL
genusDF$phylum=NULL
genusDF$class=NULL
genusDF$order=NULL
genusDF$family=NULL
rownames(genusDF) = genusDF$genus
genusDF$genus=NULL
genusDF$species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)

AvgDF1 = genusDF$C1

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$agebin == '3-5')

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$domain = NULL
genusDF$phylum=NULL
genusDF$class=NULL
genusDF$order=NULL
genusDF$family=NULL
rownames(genusDF) = genusDF$genus
genusDF$genus=NULL
genusDF$species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)

AvgDF2 = genusDF$C1

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$agebin == '6-8')

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$domain = NULL
genusDF$phylum=NULL
genusDF$class=NULL
genusDF$order=NULL
genusDF$family=NULL
rownames(genusDF) = genusDF$genus
genusDF$genus=NULL
genusDF$species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)

AvgDF3 = genusDF$C1

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$agebin == '9-11')

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$domain = NULL
genusDF$phylum=NULL
genusDF$class=NULL
genusDF$order=NULL
genusDF$family=NULL
rownames(genusDF) = genusDF$genus
genusDF$genus=NULL
genusDF$species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)

AvgDF4 = genusDF$C1

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$agebin == '12-14')

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$domain = NULL
genusDF$phylum=NULL
genusDF$class=NULL
genusDF$order=NULL
genusDF$family=NULL
rownames(genusDF) = genusDF$genus
genusDF$genus=NULL
genusDF$species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)

AvgDF5 = genusDF$C1

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$agebin == '15-17')

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$domain = NULL
genusDF$phylum=NULL
genusDF$class=NULL
genusDF$order=NULL
genusDF$family=NULL
rownames(genusDF) = genusDF$genus
genusDF$genus=NULL
genusDF$species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)

AvgDF6 = genusDF$C1

psC1 = subset_samples(ps.Genus, sample_data(ps.Genus)$agebin == '18-20')

genusDF = as.data.frame(cbind(t(otu_table(psC1, taxa_are_rows = FALSE)), tax_table(psC1)))

genusDF$domain = NULL
genusDF$phylum=NULL
genusDF$class=NULL
genusDF$order=NULL
genusDF$family=NULL
rownames(genusDF) = genusDF$genus
genusDF$genus=NULL
genusDF$species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))

genusDF$C1 = rowMeans(genusDF)

AvgDF7 = genusDF$C1 

AvgDF = cbind(AvgDF1, AvgDF2, AvgDF3, AvgDF4, AvgDF5, AvgDF6, AvgDF7) %>%
  data.frame(.)

rownames(AvgDF) = rownames(genusDF)

AvgDF$Avg = rowMeans(AvgDF)

fig = top_n(AvgDF, 9, Avg) 

Other = 1 - colSums(fig)

fig = rbind(fig, Other)

fig$tax = rownames(fig)

fig[10,9] = "Other"

fig = select(fig, -Avg) %>%
  dplyr::rename('0-2' = AvgDF1,
                '3-5' = AvgDF2,
                '6-8' = AvgDF3,
                '9-11' = AvgDF4,
                '12-14' = AvgDF5,
                '15-17' = AvgDF6,
                '18-20' = AvgDF7)

library(reshape2)
f = melt(fig) 
library(RColorBrewer)

mycolors = c("#E41A1C","#4A72A6","#48A462","#7E6E85","#EC83BA","#FFB716","#e6550d","#B75F49","#5ab4ac","#666666")

f$tax = factor(f$tax, levels=c("Anaerococcus","Corynebacterium","Dolosigranulum","Fusobacterium","Lawsonella","Moraxella", "Peptoniphilus", "Staphylococcus","Streptococcus","Other"))


ggplot(f, aes(fill = tax, y=value, x=variable)) +
  geom_bar(position="stack", stat="identity", color = "black", size = 0.25) +
  scale_fill_manual(values = mycolors, labels = c(expression(italic("Anaerococcus")),expression(italic("Corynebacterium")),expression(italic("Dolosigranulum")),expression(italic("Fusobacterium")),expression(italic("Lawsonella")),expression(italic("Moraxella")), expression(italic("Peptoniphilus")), expression(italic("Staphylococcus")),expression(italic("Streptococcus")),"Other")) +
  xlab("Age (years)") +
  ylab("Mean relative abundance") +
  labs(fill = "Genera") +
  theme_minimal() +
  scale_y_continuous(limits = c(0,1), expand = c(0, 0))+
theme(legend.text.align = 0)
```

#Figure 3A
```{r}
ancom_CorPN = read.csv("~/BRAVE_Kids/corona_ancom2_final.csv")

ancom_CorPN$ASV = as.character(ancom_CorPN$ASV)

sampleDF = sample_data(ps.NP.r) %>%
  data.frame(.)

corDF = sampleDF %>%
  select(., corona)

t=as.data.frame(ps.NP.r@tax_table@.Data[ancom_CorPN$ASV,]) %>%
  select(., 'genus')

d=otu_table(ps.NP.r)[,ancom_CorPN$ASV] %>%
  t(.) %>%
  as.data.frame(.) 
  
d1=cbind(d, t) %>%
  mutate(., var = paste(genus, " ", "(ASV", " ", rownames(.), ")", sep = "")) %>%
  select(., -genus) 
  
rownames(d1) = NULL

d=tibble::column_to_rownames(d1,var = "var") %>%
  t(.) %>%
  as.data.frame(.) %>%
  merge(., corDF, by = 0) %>%
  select(., -Row.names) %>%
  melt(., id.vars = "corona")

d$value1 = d$value + 1E-6

g=ggplot(d, aes(x = corona, y = value1, fill = corona)) +
  geom_boxplot(show.legend = F) +
  expand_limits(y=0)  +
  scale_fill_manual(values = c("darkslategray4", "cornflowerblue"))+
  theme_bw() +
  facet_wrap(~ variable, ncol=1) +
  ylab("Relative abundance") +
  xlab("SARS-CoV-2 status") +
  scale_y_continuous(breaks=c(0.000001,0.00001, 0.0001,0.001, 0.01,0.1, 1), trans = 'log') 

ggsave(file="~/BRAVE_Kids/FigurePDF/ancomPosNeg.PDF", g, width = 10, height = 16, units = "cm")
```

#Figure 3B
```{r}
ancom_CorResp = read.csv("~/BRAVE_Kids/resp_ancom2_final.csv")

ancom_age = read.csv("~/BRAVE_Kids/age_ancom2_final.csv")

#Same ASVs present, only showing those 9
ancomR = ancom_age[ancom_CorResp$ASV,]

ancom_CorResp$ASV = as.character(ancom_CorResp$ASV)

sampleDF = sample_data(ps.NP.r) %>%
  data.frame(.)

sampleDF = sampleDF %>%
    mutate(PosBin = case_when(Group == 'Negative' ~ 'Negative',
                              Group == 'PosResp' ~ 'Yes',
                              Group == 'Pos_NoResp' ~ 'No'))

RespDF = sampleDF %>%
  select(., PosBin)

ps.Pos.r = subset_samples(ps.NP.r, sample_data(ps.NP.r)$corona == "Positive")

t=as.data.frame(ps.Pos.r@tax_table@.Data[ancom_CorResp$ASV,]) %>%
  select(., 'genus')

d=otu_table(ps.Pos.r)[,ancom_CorResp$ASV] %>%
  t(.) %>%
  as.data.frame(.) 
  
d1=cbind(d, t) %>%
  mutate(., var = paste(genus, " ", "(ASV", " ", rownames(.), ")", sep = "")) %>%
  select(., -genus) 
  
rownames(d1) = NULL

d=tibble::column_to_rownames(d1,var = "var") %>%
  t(.) %>%
  as.data.frame(.) %>%
  merge(., RespDF, by = 0) %>%
  select(., -Row.names) %>%
  melt(., id.vars = "PosBin")

d$value1 = d$value + 1E-6

g=ggplot(d, aes(x = PosBin, y = value1, fill = PosBin)) +
  geom_boxplot(show.legend = F) +
  expand_limits(y=0)  +
  scale_fill_manual(values = c("mediumorchid4","#BF812D"))+
  theme_bw() +
  facet_wrap(~ variable, ncol=3) +
  ylab("Relative abundance") +
  xlab("Respiratory symptoms") +
  scale_y_continuous(breaks=c(0.000001,0.00001, 0.0001,0.001, 0.01,0.1, 1), trans = 'log')


ggsave(file="~/BRAVE_Kids/FigurePDF/ancomRespSX.PDF", g, width = 20, height = 16, units = "cm")
```

#Figure 4B
```{r}
ancom_age = read.csv("~/BRAVE_Kids/age_ancom2_final.csv")

sampleDF = sample_data(ps.NP.r) %>%
  data.frame(.)

sampleDF = sampleDF %>%
    mutate(agebin = case_when(age < 2 ~ '0-2',
                              age < 5 ~ '3-5',
                              age < 8 ~ '6-8',
                              age < 11 ~ '9-11',
                              age < 14 ~ '12-14',
                              age < 17 ~ '15-17',
                              TRUE ~ '18-20'))

sampleDF$agebin = factor(sampleDF$agebin, levels = c('0-2','3-5','6-8','9-11','12-14','15-17','18-20'))

sampleDF = sampleDF %>%
    mutate(PosBin = case_when(Group == 'Negative' ~ 'Negative',
                              Group == 'PosResp' ~ 'Yes',
                              Group == 'Pos_NoResp' ~ 'No'))

RespDF = sampleDF %>%
  select(., agebin, PosBin)

t=as.data.frame(ps.NP.r@tax_table@.Data[ancom_CorResp$ASV,]) %>%
  select(., 'genus')

d=otu_table(ps.NP.r)[,ancom_CorResp$ASV] %>%
  t(.) %>%
  as.data.frame(.) 
  
d1=cbind(d, t) %>%
  mutate(., var = paste(genus, " ", "(ASV", " ", rownames(.), ")", sep = "")) %>%
  select(., -genus) 
  
rownames(d1) = NULL

d=tibble::column_to_rownames(d1,var = "var") %>%
  t(.) %>%
  as.data.frame(.) %>%
  merge(., RespDF, by = 0) %>%
  select(., -Row.names, PosBin) %>%
  melt(., id.vars = "agebin")

d2=tibble::column_to_rownames(d1,var = "var") %>%
  t(.) %>%
  as.data.frame(.) %>%
  merge(., RespDF, by = 0) %>%
  select(., -Row.names, agebin) %>%
  melt(., id.vars = "PosBin")

d3 = merge(d, d2, by= 0) %>%
  select(., agebin, PosBin, value.x, variable.x)

d3$value.x = as.numeric(d3$value.x)

d3$value1 = d3$value.x + 1E-6

d3 = subset(d3, d3$variable.x != "PosBin" & d3$PosBin != "Negative")

g=ggplot(d3, aes(x=agebin, y=value1)) +
  geom_boxplot(aes(fill=PosBin)) +
  scale_fill_manual(values=c("mediumorchid4","#BF812D"), name = "Respiratory symptoms") +
  #stat_summary(aes(y=value1, group = GroupBin, color = GroupBin), fun.y = mean, geom = "line")+
  theme_bw()+
  xlab("Age (years)")+
  ylab("Relative abundance") + 
  
  facet_wrap(~ variable.x, ncol=3) +
  scale_y_continuous(breaks=c(0.000001,0.00001, 0.0001,0.001, 0.01,0.1, 1), trans = 'log')

ggsave(file="~/BRAVE_Kids/FigurePDF/ancomPosNeg_age.PDF", g, width = 30, height = 16, units = "cm")
```

#Figure 4A
```{r}
ancom_CorPN = read.csv("~/BRAVE_Kids/ancom2_corona_0.95_final.csv")

ancom_CorPN$ASV = as.character(ancom_CorPN$ASV)

sampleDF = sample_data(ps.NP.r) %>%
  data.frame(.)

sampleDF = sampleDF %>%
    mutate(agebin = case_when(age < 2 ~ '0-2',
                              age < 5 ~ '3-5',
                              age < 8 ~ '6-8',
                              age < 11 ~ '9-11',
                              age < 14 ~ '12-14',
                              age < 17 ~ '15-17',
                              TRUE ~ '18-20'))

sampleDF$agebin = factor(sampleDF$agebin, levels = c('0-2','3-5','6-8','9-11','12-14','15-17','18-20'))

corDF = sampleDF %>%
  select(., corona, agebin)

t=as.data.frame(ps.NP.r@tax_table@.Data[ancom_CorPN$ASV,]) %>%
  select(., 'genus')

d=otu_table(ps.NP.r)[,ancom_CorPN$ASV] %>%
  t(.) %>%
  as.data.frame(.) 
  
d1=cbind(d, t) %>%
  mutate(., var = paste(genus, " ", "(ASV", " ", rownames(.), ")", sep = "")) %>%
  select(., -genus) 
  
rownames(d1) = NULL

d=tibble::column_to_rownames(d1,var = "var") %>%
  t(.) %>%
  as.data.frame(.) %>%
  merge(., corDF, by = 0) %>%
  select(., -Row.names, agebin) %>%
  melt(., id.vars = "agebin")

d2=tibble::column_to_rownames(d1,var = "var") %>%
  t(.) %>%
  as.data.frame(.) %>%
  merge(., corDF, by = 0) %>%
  select(., -Row.names) %>%
  melt(., id.vars = "corona")

d3 = merge(d, d2, by= 0) %>%
  select(., agebin, corona, value.x, variable.x) %>%
  subset(., variable.x != "corona")

d3$value.x = as.numeric(d3$value.x)

d3$value1 = d3$value.x + 1E-6

g=ggplot(d3, aes(x = agebin, y = value1, fill = corona)) +
  geom_boxplot() +
  expand_limits(y=0)  +
  scale_fill_manual(values = c("darkslategray4", "cornflowerblue"), name = "SARS-CoV-2 status")+
  theme_bw() +
  facet_wrap(~ variable.x, ncol = 1) +
  ylab("Relative abundance") +
  xlab("Age (years)") +
  scale_y_continuous(breaks=c(0.000001,0.00001, 0.0001,0.001, 0.01,0.1, 1), trans = 'log')

ggsave(file="~/BRAVE_Kids/FigurePDF/ancomCoronaAge.PDF", g, width = 15, height = 16, units = "cm")
```

#age summary stats
```{r}
genusDF = as.data.frame(cbind(t(otu_table(ps.Genus, taxa_are_rows = FALSE)), tax_table(ps.Genus)))

genusDF$domain = NULL
genusDF$phylum=NULL
genusDF$class=NULL
genusDF$order=NULL
genusDF$family=NULL
rownames(genusDF) = genusDF$genus
genusDF$genus=NULL
genusDF$species = NULL
genusDF = mutate_all(genusDF, function(x) as.numeric(as.character(x)))


countDF = genusDF[c("Anaerococcus","Corynebacterium","Dolosigranulum","Fusobacterium","Lawsonella","Moraxella", "Peptoniphilus", "Staphylococcus","Streptococcus"),] %>%
  t(.) %>%
  data.frame(.) 

countDF$agebin = sampleDF$agebin

medDF = countDF %>%
  group_by(., agebin) %>%
  summarise_all(median) %>%
  mutate_if(is.numeric, round, digits=3)

rownames(medDF) = medDF$agebin


#######begin function 
summarystat = function(agerange){
countDF2 = countDF %>%
  subset(., agebin == agerange) %>%
  summary(.) %>%
  data.frame(.) %>%
  separate(., Freq, c("variable", "value") ,sep = ":")%>%
  select(., -Var1)  %>%
  subset(., Var2 != "   agebin") %>%
  dcast(., Var2 ~ variable)

rownames(countDF2) = countDF2$Var2
  
countDF2 = mutate_all(countDF2, as.numeric) %>%
  select(., -Var2) %>%
  mutate_if(., is.numeric, round, digits=3) 
  
countDF2$IQR = paste(countDF2$`1st Qu.`, countDF2$`3rd Qu.`, sep = "-")

countDF2$agerange = paste(countDF2$`Median `, countDF2$IQR, sep = ", ") 

countDF2 = countDF2 %>%
  select(., agerange) 
}


f1=summarystat("0-2") 
f2=summarystat("3-5") 
f3=summarystat("6-8") 
f4=summarystat("9-11")
f5=summarystat("12-14")
f6=summarystat("15-17")
f7=summarystat("18-20")

f = cbind(f1,f2,f3,f4,f5,f6,f7)

write.csv(f, "~/BRAVE_Kids/GenusSummary.csv")
```

```{r}
propcount = function(agerange){
countDF2 = countDF %>%
  subset(., agebin == agerange) %>%
  data.frame(.) %>%
  select(., -agebin) %>%
  mutate_all(function(x) ifelse(x>0, 1, 0)) %>%
  t(.) %>%
  data.frame(.)

countDF2$samples = length(countDF2)

countDF2$agebin = rowSums(countDF2) - countDF2$samples

countDF2$prop = countDF2$agebin/countDF2$samples
  
countDF2$prop = round(countDF2$prop, digits=3)

countDF2$final = paste(countDF2$agebin, countDF2$samples, sep = "/")
countDF2$final = paste("(", countDF2$final, ")", sep = "")
countDF2$final = paste(countDF2$prop, countDF2$final, sep = " ")

return(countDF2$final)
}

f1=propcount("0-2") 
f2=propcount("3-5") 
f3=propcount("6-8") 
f4=propcount("9-11")
f5=propcount("12-14")
f6=propcount("15-17")
f7=propcount("18-20")

f = cbind(f1,f2,f3,f4,f5,f6,f7) %>%
  data.frame(.)

rownames(f) = rownames(countDF2)

f = f %>%
  rename(., f1 = "0-2 years", f2 = "3-5 years", f3 = "6-8 years", f4 = "9-11 years", f5 = "12-14 years", f6 = "15-17 years", f7 = "18-20 years")

write.csv(f, "~/BRAVE_Kids/SupplementalTable3.csv")
```
